Natural selection in the Health and Retirement Study

### 2023-09-29 hrs-selection.qmd
Fixes

for (eth in eths) {
  ggp <- ggplot(tidy_pgs_joined[[eth]], aes(estimate.edyrs, estimate.fert)) +
          geom_hline(yintercept = 0, color = "grey80") +
          geom_vline(xintercept = 0, color = "grey80") +
          geom_point() +
          geom_smooth(method = "lm", formula = y ~ x, se = FALSE, 
                      linetype = "dashed", colour = "red", linewidth = 0.5) +
          scale_y_continuous(breaks = seq(-0.35, 0.15, 0.05),
                             labels = scales::label_number(0.01)) +
          coord_cartesian(xlim = c(-.3, .75)) +
          theme_light() + 
          theme(panel.grid = element_blank()) +
          labs(
            title = titles[eth],
            x = "Coefficient on years of education",
            y = "Coefficient on RLRS"
          )
  print(ggp)
}

@@

I can also examine natural selection in the previous generation, by regressing PGS on respondents' number of living siblings in 2010. (Data for dead siblings has too many missing values to use.) I reweight respondents by the reciprocal of their number of siblings, to account for parents of many siblings being more likely to be a parent of a respondent. Parents of no siblings cannot be included, so coefficient sizes are not comparable across the generations. Appendix @fig-siblings plots coefficients on number of siblings versus coefficients on years of education. Correlations are insignificant for both ethnic groups, with large standard errors (whites: correlation `r cors_sibs_edu$white[1]`, bootstrapped 95% C.I. `r ci_boot_sibs$white[1]` to `r ci_boot_sibs$white[2]`; blacks: correlation `r cors_sibs_edu$black[1]`, bootstrapped 95% C.I. `r ci_boot_sibs$black[1]` to `r ci_boot_sibs$black[2]`). However, there is a positive and significant correlation across generations, i.e. between PGS coefficients on number of siblings and number of children, of `r gen_cor`, bootstrapped 95% C.I. `r ci_gen_cor[1,1]` to `r ci_gen_cor[1,2]`.

@@

>>>>

for (eth in eths) {
  ggp <- ggplot(tidy_pgs_joined[[eth]], aes(estimate.edyrs, estimate.fert)) +
          geom_hline(yintercept = 0, color = "grey80") +
          geom_vline(xintercept = 0, color = "grey80") +
          geom_point() +
          geom_smooth(method = "lm", formula = y ~ x, se = FALSE, 
                      linetype = "dashed", colour = "red", linewidth = 0.5) +
          scale_y_continuous(breaks = seq(-0.2, 0.2, 0.02),
                             labels = scales::label_number(0.01)) +
          coord_cartesian(xlim = c(-.3, .75)) +
          theme_light() + 
          theme(panel.grid = element_blank()) +
          labs(
            title = titles[eth],
            x = "Coefficient on years of education",
            y = "Coefficient on RLRS"
          )
  print(ggp)
}

@@

I can also examine natural selection in the previous generation, by regressing PGS on respondents' number of living siblings in 2010. (Data for dead siblings has too many missing values to use.) I reweight respondents by the reciprocal of their number of siblings, to account for parents of many siblings being more likely to be a parent of a respondent. Parents of no siblings cannot be included, so coefficient sizes are not comparable across the generations. Appendix @fig-siblings plots coefficients on number of siblings versus coefficients on years of education. Correlations are insignificant for both ethnic groups, with large standard errors (whites: correlation `r cors_sibs_edu$white[1]`, bootstrapped 95% C.I. `r ci_boot_sibs$white[1]` to `r ci_boot_sibs$white[2]`; blacks: correlation `r cors_sibs_edu$black[1]`, bootstrapped 95% C.I. `r ci_boot_sibs$black[1]` to `r ci_boot_sibs$black[2]`). However, there is a positive and significant correlation across generations, i.e. between PGS coefficients on number of siblings and number of children (among whites: `r gen_cor`, bootstrapped 95% C.I. `r ci_gen_cor[1,1]` to `r ci_gen_cor[1,2]`.

@@

Code to reproduce this paper is available at <https://github.com/hughjonesd/hrs-selection>.
